### 2020-06-27 code_tests/kmeans_testing.Rmd

---
title: "k-means clustering with penguins"
author: "Allison Horst"
date: "6/26/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(GGally)
library(shadowtext)
library(NbClust)
library(patchwork)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

Following exactly along with: https://www.tidymodels.org/learn/statistics/k-means/

The "old" way, using `kmeans()`:
```{r}
penguins_k <- penguins %>% 
  filter(sex == "male") %>% 
  select(species, bill_length_mm:body_mass_g) %>% 
  drop_na()

# Using bill_length & flipper_length
# THIS ONE SHOWS *pretty* good clustering

bills <- penguins_k %>% select(species, bill_length_mm, bill_depth_mm)

set.seed(1967)
kclust_bills <- kmeans(bills[,2:3], centers = 3, nstart = 20)
kclust_bills

# Table of assigned clusters for penguins
bills_class <- table(kclust_bills$cluster,bills$species)%>% 
  as.data.frame() %>% 
  rename(cluster = Var1, species = Var2, n = Freq) %>% 
  pivot_wider(names_from = species, values_from = n)

# Same for iris:

petals <- iris %>% select(Species, Petal.Length, Petal.Width)

set.seed(1967)
kclust_petals <- kmeans(petals[,2:3], centers = 3, nstart = 25)
kclust_petals

# Table of assigned clusters for iris petals: 
petal_class <- table(kclust_petals$cluster, petals$Species) %>% 
  as.data.frame() %>% 
  rename(cluster = Var1, Species = Var2, n = Freq) %>% 
  pivot_wider(names_from = Species, values_from = n)

```

Visualizing it: 
```{r}
bills_cluster <- cbind(bills, cluster_no = kclust_bills$cluster)

penguins_kmeans <- ggplot(data = bills_cluster, aes(x = bill_length_mm, 
                                     bill_depth_mm)) +
  geom_text(aes(label=cluster_no, color = species), 
                  size = 4) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  labs(title = "Penguins k-means clustering") +
  theme(legend.position = "bottom")

iris_cluster <-cbind(petals, cluster_no = kclust_petals$cluster)

iris_kmeans <- ggplot(data = iris_cluster, aes(x = Petal.Length, y = Petal.Width)) +
  geom_text(aes(label = cluster_no, color = Species), size = 4) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  theme(panel.background = element_rect(fill = "gray90", color = "gray90"),
        legend.position = "bottom") +
  labs(title = "Iris k-means clustering")

(penguins_kmeans | iris_kmeans)

# Full table of cluster assignments: 
combo_cluster <- cbind(bills_class, petal_class) %>% 
  kable() %>% 
  kable_styling(full_width = FALSE)

combo_cluster

```
## Using tidymodels? eeeeehhh we'll see