### 2021-04-27 penguins.tex

% !TeX root = RJwrapper.tex
\title{Palmer Archipelago Penguins Data in the palmerpenguins R Package
- An Alternative to Anderson's Irises}
\author{by Allison M. Horst, Alison Presmanes Hill, and Kristen B. Gorman}

\maketitle

\abstract{%
In 1935, Edgar Anderson collected size measurements for 150 flowers from
three species of \emph{Iris} on the Gaspé Peninsula in Quebec, Canada.
Since then, Anderson's \emph{Iris} observations have become a classic
dataset in statistics, machine learning, and data science teaching
materials. It is included in the base R \textbf{datasets} package as
\textbf{iris}, making it easy for users to access without knowing much
about it. However, the lack of data documentation, presence of
non-intuitive variables (e.g.~``sepal width''), and perfectly balanced
groups with zero missing values make \textbf{iris} an inadequate and
stale dataset for teaching and learning modern data science skills.
Users would benefit from working with a more representative, real-world
environmental dataset with a clear link to current scientific research.
Importantly, Anderson's \emph{Iris} dataset appeared in a 1936
publication by R. A. Fisher in the \emph{Annals of Eugenics} (which is
often the first-listed citation for the dataset), inextricably linking
\textbf{iris} to racist research and theory. Thus, a modern alternative
to the \textbf{iris} dataset is needed. In this paper, we introduce the
\textbf{palmerpenguins} R package, which includes morphological
measurements collected from 2007 - 2009 for three species of
\emph{Pygoscelis} penguins that breed on islands throughout the Palmer
Archipelago, Antarctica. The \textbf{penguins} data provides an
approachable, charismatic, and near drop-in replacement for
\textbf{iris} with topical relevance global climate change and
environmental impacts on marine predators. Since the release on CRAN in
July 2020, the \textbf{palmerpenguins} package has been downloaded over
102,000 times, highlighting the demand and widespread adoption of this
viable \textbf{iris} alternative. We directly compare the \textbf{iris}
and \textbf{penguins} datasets for selected analyses to demonstrate that
R users, in particular teachers and learners currently using
\textbf{iris}, can switch to the Palmer Archipelago penguins for many
use cases including data wrangling, visualization, linear modeling,
multivariate analysis (e.g., PCA), cluster analysis and classification
(e.g., by k-means).
}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

In 1935, American botanist Edgar Anderson measured petal and sepal
structural dimensions (length and width) for 50 flowers from three
\emph{Iris} species: \emph{Iris setosa}, \emph{Iris versicolor}, and
\emph{Iris virginica} \citep{anderson_irises_1935}. The manageable but
non-trivial size (5 variables and 150 total observations) and
characteristics of the dataset, including linear relationships and
multivariate normality, have made the dataset amenable for introducing a
wide range of statistical methods including data wrangling,
visualization, linear modeling, multivariate analyses, and machine
learning. Anderson's \emph{Iris} data is built into a number of software
packages including the auto-installed datasets package in R
\citep{r_core_team_r_2019}, Python's scikit-learn machine learning
library \citep{pedregosa_scikit-learn_2011}, and the SAS Sashelp library
(SAS Institute, Cary NC), which have facilitated widespread use of the
\emph{Iris} data. As a result, eighty-six years after the data were
initially published, Anderson's \emph{Iris} measurements are ubiquitous
in statistics, computational methods, software documentation, and data
science courses and materials.

There are a number of reasons that modern data science practitioners and
educators may want to move on from Anderson's \emph{Iris} dataset.
First, Anderson's \emph{Iris} dataset lacks metadata almost entirely
\citep{anderson_irises_1935}, which does not reinforce best practices
and limits meaningful interpretation and discussion of research methods,
analyses, and outcomes. Of the five variables in the dataset, two
(\texttt{Sepal.Width} and \texttt{Sepal.Length}) are not intuitive for
most non-botanists. Even with explanation, the difference between
\emph{Iris} \emph{petal} and \emph{sepal} dimensions is not obvious.
Second, the \emph{Iris} dataset contains equal sample sizes for each of
the three species (\emph{n} = 50) with no missing values, which is
cleaner than most real-world datasets that learners are likely to
encounter ``in the wild.'' Third, the single factor (\texttt{Species})
in the \emph{Iris} data limits options for analyses. Finally, due to its
publication in the \emph{Annals of Eugenics} by eugenicist and
statistician R.A. Fisher \citep{fisher_use_1936}, the \emph{Iris}
dataset is burdened by a history in scientific racism, which we are
committed to addressing through the development of new data science
education products as described below.

Given the growing need for fresh data science-ready datasets, we sought
to identify an alternative dataset that could be made easily accessible
for a broad audience. After evaluating the positive and negative
features of the \emph{Iris} dataset in data science and statistics
materials, we established the following criteria for a suitable
alternative:

\begin{itemize}
\tightlist
\item
  Available by appropriate license (ideally, CC0 ``no rights reserved'')
\item
  Feature intuitive subjects and variables that are interesting and
  understandable to learners across disciplines
\item
  Complete metadata and documentation
\item
  Manageable (but not trivial) in size
\item
  Minimal data cleaning and pre-processing required for most analyses
\item
  Real-world (not manufactured) modern data
\item
  Provides similar opportunities for teaching and learning R, data
  science, and statistical skills
\item
  Can easily replace \emph{Iris} for most use cases
\end{itemize}

Here, we describe an alternative to Anderson's \emph{Iris} data that
largely satisfies these criteria: a refreshing, approachable, and
charismatic dataset containing real-world morphological data for three
\emph{Pygoscelis} penguin species that breed throughout the Western
Antarctic Peninsula region, made available through the United States
Long-Term Ecological Research (US LTER) Network. By comparing data
structure, size, and a range of analyses side-by-side for the two
datasets, we demonstrate that the Palmer Archipelago penguin data are an
ideal substitute for Anderson's \emph{Iris} dataset for many use cases
in statistics and data science education.

\hypertarget{data-source}{%
\subsection{Data source}\label{data-source}}

Body size measurements, clutch (i.e., egg laying) observations (e.g.,
date of first egg laid, and clutch completion), and carbon
(\textsuperscript{13}C/\textsuperscript{12}C,
\(\delta\)\textsuperscript{13}C) and nitrogen
(\textsuperscript{15}N/\textsuperscript{14}N,
\(\delta\)\textsuperscript{15}N) stable isotope values of red blood
cells for adult male and female Adélie (\emph{P. adeliae}), chinstrap
(\emph{P. antarcticus}), and gentoo (\emph{P. papua}) penguins on three
islands (Biscoe, Dream and Torgersen) within the Palmer Archipelago were
collected from 2007 - 2009 by Dr.~Kristen Gorman in collaboration with
the \href{https://pal.lternet.edu/}{Palmer Station LTER}, part of the
\href{https://lternet.edu/}{US LTER Network}. For complete data
collection methods and published analyses, see
\citet{gorman_ecological_2014}. Throughout this paper, penguins species
referred to as ``Adélie'', ``Chinstrap'', and ``Gentoo.''

The data in the \CRANpkg{palmerpenguins} R package are available for use
by CC0 license (``No Rights Reserved'') in accordance with the
\href{https://pal.lternet.edu/data/policies}{Palmer Station LTER Data
Policy} and the \href{https://lternet.edu/data-access-policy/}{LTER Data
Access Policy}, and were imported from the
\href{https://environmentaldatainitiative.org/}{Environmental Data
Initiative (EDI) Data Portal} at the links below:

\begin{itemize}
\tightlist
\item
  Adélie penguin data (LTER and Gorman 2020a):
  \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-pal.219.5}{KNB-LTER
  Data Package 219.5}
\item
  Chinstrap penguin data (LTER and Gorman 2020b):
  \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-pal.221.6}{KNB-LTER
  Data Package 221.5}
\item
  Gentoo penguin data (LTER and Gorman 2020c):
  \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-pal.220.5}{KNB-LTER
  Data Package 220.5}
\end{itemize}

\hypertarget{r-package-palmerpenguins}{%
\subsection{R package: palmerpenguins}\label{r-package-palmerpenguins}}

R users can install the \CRANpkg{palmerpenguins} package from CRAN:

\begin{verbatim}
install.packages("palmerpenguins")
\end{verbatim}

Information, examples, and links to community-contributed materials are
available on the \CRANpkg{palmerpenguins} package website:
\url{https://allisonhorst.github.io/palmerpenguins/}. See the Appendix
for how Python and Julia users can access the same data.

The \CRANpkg{palmerpenguins} R package contains two data objects:
\textbf{penguins\_raw} and \textbf{penguins}. The \textbf{penguins\_raw}
data consists of all raw data for 17 variables, recorded completely or
in part for 344 individual penguins, accessed directly from EDI
(\textbf{penguins\_raw} properties are summarized in Appendix B). We
generally recommend using the curated data in \textbf{penguins}, which
is a subset of \textbf{penguins\_raw} retaining all 344 observations,
minimally updated (Appendix A) and reduced to the following eight
variables:

\begin{itemize}
\tightlist
\item
  \emph{species:} a factor denoting the penguin species (Adélie,
  Chinstrap, or Gentoo)
\item
  \emph{island:} a factor denoting the Palmer Archipelago island in
  Antarctica where each penguin was observed (Biscoe Point, Dream
  Island, or Torgersen Island)
\item
  \emph{bill\_length\_mm:} a number denoting length of the dorsal ridge
  of a penguin bill (millimeters)
\item
  \emph{bill\_depth\_mm:} a number denoting the depth of a penguin bill
  (millimeters)
\item
  \emph{flipper\_length\_mm:} an integer denoting the length of a
  penguin flipper (millimeters)
\item
  \emph{body\_mass\_g:} an integer denoting the weight of a penguin's
  body (grams)
\item
  \emph{sex:} a factor denoting the sex of a penguin sex (male, female)
  based on molecular data
\item
  \emph{year:} an integer denoting the year of study (2007, 2008, or
  2009)
\end{itemize}

\newpage

The same data exist as comma-separated value (CSV) files in the package
(``penguins\_raw.csv'' and ``penguins.csv''), and can be read in using
the built-in \texttt{path\_to\_file()} function in
\CRANpkg{palmerpenguins}. For example,

\begin{verbatim}
    library(palmerpenguins)
    df <- read.csv(path_to_file("penguins.csv"))
\end{verbatim}

will read in ``penguins.csv'' as if from an external file, thus
automatically parsing \emph{species}, \emph{island}, and \emph{sex}
variables as characters instead of factors. This option allows users
opportunities to practice or demonstrate reading in data from a CSV,
then updating variable class (e.g., characters to factors).

\hypertarget{comparing-iris-and-penguins}{%
\subsection{\texorpdfstring{Comparing \textbf{iris} and
\textbf{penguins}}{Comparing iris and penguins}}\label{comparing-iris-and-penguins}}

The \textbf{penguins} data in \CRANpkg{palmerpenguins} is useful and
approachable for data science and statistics education, and is uniquely
well-suited to replace the \textbf{iris} dataset. Comparisons presented
are selected examples for common \textbf{iris} uses, and are not
exhaustive.

\begin{Schunk}
\begin{table}

\caption{\label{tab:overview-tbl}Overview comparison of penguins and iris dataset features and characteristics.}
\centering
\begin{tabular}[t]{lcl}
\toprule
Feature & iris & penguins\\
\midrule
Year(s) collected & 1935 & 2007 - 2009\\
Dimensions (col x row) & 5 x 150 & 8 x 344\\
Documentation & minimal & complete metadata\\
Variable classes & double (4), factor (1) & double (2), int (3), factor (3)\\
Missing values? & no (n = 0; 0.0\%) & yes (n = 19; 0.7\%)\\
\bottomrule
\end{tabular}
\end{table}

\end{Schunk}

\hypertarget{data-structure-and-sample-size}{%
\subsubsection{Data structure and sample
size}\label{data-structure-and-sample-size}}

Both \textbf{iris} and \textbf{penguins} are in tidy format
\citep{wickham_tidy_2014} with each column denoting a single variable
and each row containing measurements for a single \emph{Iris} flower or
penguin, respectively. The two datasets are comparable in size:
dimensions (columns × rows) are 5 × 150 and 8 × 344 for \textbf{iris}
and \textbf{penguins}, respectively, and sample sizes within species are
similar (Table 1).

Notably, while sample sizes in \textbf{iris} across species are all the
same, sample sizes in \textbf{penguins} differ across the three species.
The inclusion of three factor variables in \textbf{penguins}
(\emph{species}, \emph{island}, and \emph{sex}), along with \emph{year},
create additional opportunities for grouping, faceting, and analysis
compared to the single factor (\emph{Species}) in \textbf{iris}.

Unlike \textbf{iris}, which contains only complete cases, the
\textbf{penguins} dataset contains a small number of missing values
(n\textsubscript{missing} = 19, out of 2752 total values). Missing
values and unequal sample sizes are common in real-world data, and adds
learning opportunities to the \textbf{penguins} dataset.

\begin{Schunk}
\begin{table}

\caption{\label{tab:counts-tbl}Grouped sample size for iris (by species; n = 150 total) and penguins (by species and sex; n = 344 total). Penguins can be further grouped by variables for island and study year.}
\centering
\begin{tabular}[t]{lclccc}
\toprule
\multicolumn{2}{c}{Iris sample size (by species)} & \multicolumn{4}{c}{Penguin sample size (by species and sex)} \\
\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-6}
Iris species & Sample size & Penguin species & Female & Male & NA\\
\midrule
setosa & 50 & Adélie & 73 & 73 & 6\\
versicolor & 50 & Chinstrap & 34 & 34 & 0\\
virginica & 50 & Gentoo & 58 & 61 & 5\\
\bottomrule
\end{tabular}
\end{table}

\end{Schunk}

\hypertarget{continuous-quantitative-variables}{%
\subsubsection{Continuous quantitative
variables}\label{continuous-quantitative-variables}}

Distributions, relationships between variables, and clustering can be
visually explored between species for the four structural size
measurements in \textbf{penguins} (flipper length, body mass, bill
length and depth; Figure \ref{fig:penguin-pairs}) and \textbf{iris}
(sepal width and length, petal width and length; Figure
\ref{fig:iris-pairs}).

\begin{Schunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figs/penguin-pairs-1}

}

\caption[Distributions and correlations for numeric variables in the penguins data (flipper length (mm), body mass (g), bill length (mm) and bill depth (mm)) for the three observed species]{Distributions and correlations for numeric variables in the penguins data (flipper length (mm), body mass (g), bill length (mm) and bill depth (mm)) for the three observed species: Gentoo (green, triangles); Chinstrap (blue, circles); and Adélie (orange, squares). Significance levels indicated for bivariate correlations are: *p < 0.05; **p < 0.01; ***p < 0.001.}\label{fig:penguin-pairs}
\end{figure}
\end{Schunk}

\begin{Schunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figs/iris-pairs-1}

}

\caption[Distributions and correlations for numeric variables in the iris data (petal length (cm), petal width (cm), sepal length (cm) and sepal width (cm)) for the three included Iris species]{Distributions and correlations for numeric variables in the iris data (petal length (cm), petal width (cm), sepal length (cm) and sepal width (cm)) for the three included Iris species: setosa (light gray, circles); versicolor (dark gray, triangles); and virginica (black, squares). Significance levels indicated for bivariate correlations are: *p < 0.05; **p < 0.01; ***p < 0.001.}\label{fig:iris-pairs}
\end{figure}
\end{Schunk}

Both \textbf{penguins} and \textbf{iris} offer numerous opportunities to
explore linear relationships and correlations, within and across species
(Figures \ref{fig:penguin-pairs} \& \ref{fig:iris-pairs}). A bivariate
scatterplot made with the \textbf{iris} dataset reveals a clear linear
relationship between petal length and petal width. Using
\textbf{penguins} (\ref{fig:linear-example}), we can create a uniquely
similar scatterplot with flipper length and body mass. The overall trend
across all three species is approximately linear for both \textbf{iris}
and \textbf{penguins}. Teachers may encourage students to explore how
simple linear regression results and predictions differ when the species
variable is omitted, compared to multiple linear regression with species
included (\ref{fig:linear-example}).

\begin{Schunk}
\begin{figure}[htbp]

{\centering \includegraphics[width=1\linewidth]{figs/linear-example-1}

}

\caption[Representative linear relationships for (A)]{Representative linear relationships for (A): penguin flipper length (mm) and body mass (g) for Adélie (orange circles), Chinstrap (blue triangles), and Gentoo (green squares) penguins; (B): iris petal length (cm) and width (cm) for setosa (light gray circles), versicolor (dark gray triangles) and virginica (black squares) irises. Within-species linear model is visualized for each penguin or iris species.}\label{fig:linear-example}
\end{figure}
\end{Schunk}

Notably, distinctions between species are clearer for iris petals -
particularly, the much smaller petals for \emph{Iris setosa} - compared
to penguins, in which Adélie and Chinstrap penguins are largely
overlapping in body size (body mass and flipper length), and are both
generally smaller than Gentoo penguins.

Simpson's Paradox is a data phenomenon in which a trend observed between
variables is reversed when data are pooled, omitting a meaningful
variable. While often taught and discussed in statistics courses,
finding a real-world and approachable example of Simpson's Paradox can
be a challenge. Here, we show one (of several possible - see Figure
\ref{fig:penguin-pairs}) Simpson's Paradox examples in
\textbf{penguins}: exploring bill dimensions with and without species
included (Figure \ref{fig:simpsons}). When penguin species is omitted
(Figure \ref{fig:simpsons} A), bill length and depth appear negatively
correlated overall. The trend is reversed when species is included,
revealing an obviously positive correlation between bill length and bill
depth within species (Figure \ref{fig:simpsons} B).

\begin{Schunk}
\begin{figure}[htbp]

{\centering \includegraphics[width=1\linewidth]{figs/simpsons-1}

}

\caption[Trends for penguin bill dimensions (bill length and bill depth, millimeters) if the species variable is excluded (A) or included (B), illustrating Simpson’s Paradox]{Trends for penguin bill dimensions (bill length and bill depth, millimeters) if the species variable is excluded (A) or included (B), illustrating Simpson’s Paradox. Note: linear regression for bill dimensions without including species in (A) is ill-advised; the linear trendline is only included to visualize trend reversal for Simpson’s Paradox when compared to (B).}\label{fig:simpsons}
\end{figure}
\end{Schunk}

\hypertarget{principal-component-analysis}{%
\subsubsection{Principal component
analysis}\label{principal-component-analysis}}

Principal component analysis (PCA) is a dimensional reduction method
commonly used to explore patterns in multivariate data. The
\textbf{iris} dataset frequently appears in PCA tutorials due to
multivariate normality and clear interpretation of variable loadings and
clustering.

A comparison of PCA with the four variables of structural size
measurements in \textbf{penguins} and \textbf{iris} (both normalized
prior to PCA) reveals highly similar results (Figure \ref{fig:pca}). For
both datasets, one species is distinct (Gentoo penguins, and setosa
irises) while the other two species (Chinstrap/Adélie and
versicolor/virginica) appear somewhat overlapping in the first two
principal components (Figure \ref{fig:pca} A,B). Screeplots reveal that
the variance explained by each principal component (PC) is very similar
across the two datasets, particularly for PC1 and PC2: for
\textbf{penguins}, 88.15\% of total variance is captured by the first
two PCs, compared to 95.81\% for \textbf{iris}, with a similarly large
percentage of variance captured by PC1 and PC2 in each (Figure
\ref{fig:pca} C,D).

\begin{Schunk}
\begin{figure}[htbp]

{\centering \includegraphics[width=1\linewidth]{figs/pca-1}

}

\caption[Principal component analysis biplots and screeplots for structural size measurements in penguins (A,C) and iris (B,D), revealing similarities in multivariate patterns, variable loadings, and variance explained by each component]{Principal component analysis biplots and screeplots for structural size measurements in penguins (A,C) and iris (B,D), revealing similarities in multivariate patterns, variable loadings, and variance explained by each component. For penguins, variables are flipper length (mm), body mass (g), bill length (mm) and bill depth (mm); groups are visualized by species (Adélie = orange circles, Chinstrap = blue triangles, Gentoo = green squares). For iris, variables are petal length (cm), petal width (cm), sepal length (cm) and sepal width (cm); groups are visualized by species (setosa = light gray circles, versicolor = dark gray triangles, virginica = black squares). Values above screeplot columns (C,D) indicate percent of total variance explained by each of the four principal components.}\label{fig:pca}
\end{figure}
\end{Schunk}

\hypertarget{k-means-clustering}{%
\subsubsection{K-means clustering}\label{k-means-clustering}}

Unsupervised clustering by k-means is a common and popular entryway to
machine learning and classification, and again, the \textbf{iris}
dataset is frequently used in introductory examples. The
\textbf{penguins} data provides similar opportunities for introducing
k-means clustering. For simplicity, we compare k-means clustering using
only two variables for each dataset: for \textbf{iris}, petal width and
petal length, and for \textbf{penguins}, bill length and bill depth. All
variables are scaled prior to k-means. Three clusters (\emph{k} = 3) are
specified for each, since there are three species of both \emph{Iris}
(\emph{setosa}, \emph{versicolor}, and \emph{virginica}) and penguins
(Adélie, Chinstrap and Gentoo).

K-means clustering with penguin bill dimensions and iris petal
dimensions yields largely distinct clusters, each dominated by one
species (\ref{fig:kmeans}). For iris petal dimensions, k-means yields a
perfectly separated cluster (Cluster 3) containing all 50 setosa iris
observations and zero misclassified virginica or versicolor irises
(Table 2). While clustering is not perfectly distinct for any penguin
species, each species is largely contained within a single cluster, with
little overlap from the other two species. For example, considering
Adélie penguins (orange observations in Figure \ref{fig:kmeans} A): 147
(out of 151) Adélie penguins are assigned to Cluster 3, zero are
assigned to Cluster 1, and 4 are assigned to the Chinstrap-dominated
Cluster 2 (Table 2). Only 5 (of 68) Chinstrap penguins and 1 (of 123)
Gentoo penguins are assigned to the Adélie-dominated Cluster 3 (Table
2).

\begin{Schunk}
\begin{figure}[htbp]

{\centering \includegraphics[width=1\linewidth]{figs/kmeans-1}

}

\caption[K-means clustering outcomes for penguin bill dimensions (A) and iris petal dimensions (B)]{K-means clustering outcomes for penguin bill dimensions (A) and iris petal dimensions (B). Numbers indicate the cluster to which an observation was assigned, revealing a high degree of separation between species for both penguins and iris.}\label{fig:kmeans}
\end{figure}
\end{Schunk}

\begin{Schunk}
\begin{table}

\caption{\label{tab:unnamed-chunk-1}K-means cluster assignments by species based on penguin bill length (mm) and depth (mm), and iris petal length (cm) and width (cm).}
\centering
\begin{tabular}[t]{cccccccc}
\toprule
\multicolumn{4}{c}{Penguins cluster assignments} & \multicolumn{4}{c}{Iris cluster assignments} \\
\cmidrule(l{3pt}r{3pt}){1-4} \cmidrule(l{3pt}r{3pt}){5-8}
Cluster & Adélie & Chinstrap & Gentoo & Cluster & setosa & versicolor & virginica\\
\midrule
1 & 0 & 9 & 116 & 1 & 0 & 2 & 46\\
2 & 4 & 54 & 6 & 2 & 0 & 48 & 4\\
3 & 147 & 5 & 1 & 3 & 50 & 0 & 0\\
\bottomrule
\end{tabular}
\end{table}

\end{Schunk}

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

Here, we have shown that structural size measurements for Palmer
Archipelago \emph{Pygoscelis} penguins, available as \textbf{penguins}
in the \CRANpkg{palmerpenguins} R package, offer a near drop-in
replacement for \textbf{iris} in a number of common use cases for data
science and statistics education including exploratory data
visualization, linear correlation and regression, PCA, and clustering by
k-means. In addition, teaching and learning opportunities in
\textbf{penguins} are increased due to a greater number of variables,
missing values, unequal sample sizes, and Simpson's paradox examples.
Importantly, the \textbf{penguins} dataset encompasses real-world
information derived from several charismatic marine predator species
with regional breeding populations notably responding to environmental
change occurring throughout the Western Antarctic Peninsula region of
the Southern Ocean (see \citet{bestelmeyer_analysis_2011},
\citet{gorman_ecological_2014}, \citet{gorman_integrative_2015},
\citet{gorman_population_2017}). Thus, the \textbf{penguins} dataset can
facilitate discussions more broadly on biodiversity responses to global
change - a contemporary and critical topic in ecology, evolution, and
the environmental sciences.

\hypertarget{appendix}{%
\section*{Appendix}\label{appendix}}
\addcontentsline{toc}{section}{Appendix}

\hypertarget{penguins-data-processing}{%
\subsection{Penguins data processing}\label{penguins-data-processing}}

Data in the \textbf{penguins} object have been minimally updated from
\textbf{penguins\_raw} as follows:

\begin{itemize}
\tightlist
\item
  All variable names are converted to lower snake case (e.g.~from
  \texttt{Flipper\ Length\ (mm)} to \texttt{flipper\_length\_mm})
\item
  Entries in \emph{species} are truncated to only include the common
  name (e.g.~``Gentoo'', instead of ``gentoo penguin (Pygoscelis
  papua)'')
\item
  Recorded sex for penguin N36A1, originally recorded as ``.'', is
  updated to \texttt{NA}
\item
  \emph{culmen\_length\_mm} and \emph{culmen\_depth\_mm} variable names
  are updated to \emph{bill\_length\_mm} and \emph{bill\_depth\_mm},
  respectively
\item
  Class for categorical variables (\emph{species}, \emph{island},
  \emph{sex}) is updated to factor
\item
  Variable \emph{year} was pulled from clutch observations
\end{itemize}

\hypertarget{summary-of-the-penguins_raw-dataset}{%
\subsection{\texorpdfstring{Summary of the \textbf{penguins\_raw}
dataset}{Summary of the penguins\_raw dataset}}\label{summary-of-the-penguins_raw-dataset}}

\begin{Schunk}

\begin{tabular}{ll}
\toprule
Feature & penguins\_raw\\
\midrule
Year(s) collected & 2007 - 2009\\
Dimensions (col x row) & 17 x 344\\
Documentation & complete metadata\\
Variable classes & character (9), Date (1), numeric (7)\\
Missing values? & yes (n = 336; 5.7\%)\\
\bottomrule
\end{tabular}

\end{Schunk}

\hypertarget{other-data-sources}{%
\subsection{Other data sources}\label{other-data-sources}}

Python: Python users can access the penguins data in the
\textbf{seaborn} data visualization library
\citep{michael_waskom_2017_883859}. Example code to load the data in
Python:

\begin{verbatim}
    import seaborn as sns
    df = sns.load_dataset('penguins') 
\end{verbatim}

Julia: Julia users can access the penguins data in the
\textbf{PalmerPenguins.jl} package. Example code to import the penguins
data through \textbf{PalmerPenguins.jl} (more information on
\textbf{PalmerPenguins.jl} from David Widmann can be found
\href{https://github.com/devmotion/PalmerPenguins.jl}{here}):

\begin{verbatim}
    julia> using PalmerPenguins
    julia> table = PalmerPenguins.load()
\end{verbatim}

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

All analyses were performed in the R language environment using version
3.6.2 \citep{r_core_team_r_2019}. Complete code for this paper is shared
in the Supplemental Material. We acknowledge the following R packages
used in analyses, with gratitude to developers and contributors:

\begin{itemize}
\tightlist
\item
  \CRANpkg{tidyverse} \citep{wickham_welcome_2019}: for data import and
  cleaning
\item
  \CRANpkg{ggplot2} \citep{wickham_ggplot2_2016}: for data
  visualizations
\item
  \CRANpkg{kableExtra} \citep{zhu_kableextra_2019}: for finalized tables
\item
  \CRANpkg{GGally} \citep{schloerke_ggally_2020}: for pairs plots
\item
  \CRANpkg{patchwork} \citep{pedersen_patchwork_2019}: for compound
  figures
\item
  \CRANpkg{shadowtext} \citep{yu_shadowtext_2019}: to add a background
  color to text labels
\item
  \CRANpkg{recipes} \citep{kuhn_recipes_2020}: for data pre-processing
\item
  \CRANpkg{pkgdown} \citep{wickham_pkgdown_2020}: to build the package
  website
\item
  \pkg{base} and \pkg{stats} \citep{r_core_team_r_2019}: for various
  analyses throughout
\end{itemize}

\bibliography{RJreferences}

\address{%
Allison M. Horst\\
Bren School of Environmental Science and Management\\%
University of California, Santa Barbara\\ Santa Barbara, CA 93106-5131\\
%
%
%
\\\href{mailto:ahorst@ucsb.edu}{\nolinkurl{ahorst@ucsb.edu}}
}

\address{%
Alison Presmanes Hill\\
RStudio, PBC\\%
250 Northern Ave\\ Boston, MA 02210\\
%
%
%
\\\href{mailto:alison@rstudio.com}{\nolinkurl{alison@rstudio.com}}
}

\address{%
Kristen B. Gorman\\
College of Fisheries and Ocean Sciences\\%
University of Alaska Fairbanks\\ 2150 Koyukuk Drive\\ 245 O'Neill
Building\\ Fairbanks, AK 99775-7220\\
%
%
%
\\\href{mailto:kbgorman@alaska.edu}{\nolinkurl{kbgorman@alaska.edu}}
}